---
title: "Using fdim"
author: "Pramit Ghosh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using fdim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

This vignette aims at providing an overview of how the calculation of the fractal dimension using the box-counting technique is carried out and then proceeds to give an example.

## Calculating the Box-Counting Dimension

The fdim package is used for calculating the fractal dimension of a simple feature of geometry type `POLYGON` using the box-counting technique. It is done by performing the following important steps:

- Importing a simple feature as an object of class `sf`.
- If the imported simple feature has spherical coordinates ($\mathbb{S}^2$), it is projected conformally using the Mercator projection (EPSG:3857) so that it is referenced using Cartesian coordinate reference system ($\mathbb{R}^2$).
- A grid of varying grid sizes in the same CRS as the simple feature is overlayed on the feature. The default sequence of cell cizes are 
```{r}
seq(10000, 100000, 10000)
```

- The number of cells intersecting the feature is counted for grids of each cell size and recorded against the inverse of the cell size.
- The count of the number of boxes (log-transformed) is plotted against the reciprocal of the cell-size (log-transformed).
- A linear regression is performed to find the best-fit line for this log-log scatterplot.
- The slope of the best-fit line gives the box-counting dimension for the feature.

## Example

```{r}
library('fdim')
```

`fdim` allows the import of a simple feature in two ways.

- If the feature is stored in an external file, it may be imported using the `dsn` and `layer` arguments in the function `bcd()` along with any other arguments for `sf::st_read()`. The `plot = TRUE` argument plots the scatterplot and the shows the best-fit line obtained from the simple linear regression. The default argument for `plot` is `FALSE`.

```{r fig.height=5, fig.width=7}
bcd(dsn = system.file(package = "fdim"), layer = "madhya_pradesh", plot = TRUE)
```

- Alternatively, if the simple feature is already available to the user in the envrionment, it can be passed on to `bcd()` as the the argument `x`. The sequence of cell-sizes to be tested can be optionally modified by explicitly passing the argument `l` to `bcd()`.

```{r fig.height=5, fig.width=7}
library('rnaturalearth')
deutschland = ne_countries(scale = "medium", country = "Germany", returnclass = "sf")
bcd(deutschland, l = seq(10000, 100000, 15000), plot = TRUE)
```





