---
title: "Using sameSVD"
author: "Pramit Ghosh"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using sameSVD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

This vignette aims at providing an overview of how the calculation of the fractal dimension using the box-counting technique is carried out and then proceeds to give an example.

# Calculating the Box-Counting Dimension

The `sameSVD` package is used for calculating the fractal dimension of a simple feature of geometry type `POLYGON` using the box-counting technique. It is done by performing the following important steps:

- Importing a simple feature as an object of class `sf`.
- If the imported simple feature has spherical coordinates ($\mathbb{S}^2$), it is projected conformally using the Mercator projection (EPSG:3857) so that it is referenced using Cartesian coordinate reference system ($\mathbb{R}^2$).
- A grid of varying grid sizes in the same CRS as the simple feature is overlayed on the feature. The default sequence of cell cizes are 
```{r}
seq(10000, 100000, 10000)
```

- The number of cells intersecting the feature is counted for grids of each cell size and recorded against the inverse of the cell size.
- The count of the number of boxes (log-transformed) is plotted against the reciprocal of the cell-size (log-transformed).
- A linear regression is performed to find the best-fit line for this log-log scatterplot.
- The slope of the best-fit line gives the box-counting dimension for the feature.

# Example

```{r}
library('sameSVD')
```

`sameSVD` allows the import of a simple feature in two ways.

- If the feature is stored in an external file, it may be imported using the `dsn` and `layer` arguments in the function `bcd()` along with any other arguments for `sf::st_read()`. The `plot = TRUE` argument plots the scatterplot and the shows the best-fit line obtained from the simple linear regression. The default argument for `plot` is `FALSE`.

```{r fig.height=5, fig.width=7}
mp = import_SVD(dsn = system.file(package = "sameSVD"), layer = "madhya_pradesh")
bcd(k = mp, l = seq(50000, 100000, 10000), plot = TRUE)
```

- Alternatively, if the simple feature is already available to the user in the envrionment, it can be passed on to `bcd()` as the the argument `x`. The sequence of cell-sizes to be tested can be optionally modified by explicitly passing the argument `l` to `bcd()`.

```{r fig.height=5, fig.width=7}
library('rnaturalearth')
deutschland = import_SVD(ne_countries(scale = "medium", country = "Germany", returnclass = "sf"))
bcd(deutschland, l = seq(10000, 100000, 15000), plot = TRUE)
```

## Internal working

When the above code is executed, `deutschland` stores a `sf` polygon representing Germany. When `bcd()` is called it creates `sf` objects in the form of grids (comprising of polygons) covering `deutschland`. Additionally, since the CRS of `deutschland` is in latitude and longitude, it transforms it to EPSG:3857.

```{r}
library(sf)
deutschland = st_transform(deutschland, 3857)
```

The number of grids created is the length of the vector `l` that is passed to `bcd()`. Each grid is made up of square polygons with side-lengths `l`. The following figures show square grids overlayed on top of `deutschland` with side-lengths 40000m and 70000m, as an illustraion.

```{r fig.height=5, fig.width=7}
par(mfrow=c(1,2))

plot(st_geometry(deutschland), col = "wheat", axes = TRUE, main = "Cell size = 40000m", xlab = "Easting (m)", ylab = "Northing (m)")
plot(st_make_grid(deutschland, cellsize = 40000), add = TRUE)

plot(st_geometry(deutschland), col = "wheat", axes = TRUE, main = "Cell size = 70000m", xlab = "Easting (m)", ylab = "Northing (m)")
plot(st_make_grid(deutschland, cellsize = 70000), add = TRUE)

par(mfrow = c(1,1))
```

The number of cells intersecting with `deutschland`, say $N$, is counted for each grid with cells having side-length $l$. A simple linear regression is performed with $log(N)$ being the dependent variable and $-log(\frac{1}{l})$. The coefficient of the dependent variable i.e., the slope of the best-fit line gives the Box-Counting Dimension.

In the above example, the number of intersecting cells for cell sizes 40000m and 70000m are
```{r}
# For cell size = 40000m
(n1 = length(st_intersection(st_geometry(deutschland), st_make_grid(deutschland, cellsize = 40000))))
# For cell size = 70000m
(n2 = length(st_intersection(st_geometry(deutschland), st_make_grid(deutschland, cellsize = 70000))))
```
respectively. Thus, the dependent variable $N$ would be
```{r}
(y = log(c(n1, n2)))
```
while the independent variable $-log(\frac{1}{l})$ would be
```{r}
(x = -log(c(1/40000, 1/70000)))
```

The slope of the best-fit line in this case will be the slope of the line joining the points and is equal to
```{r}
-(y[2] - y[1])/(x[2] - x[1])
```

This is very close to the slope calculated using `bcd()` earlier in this document.